
---
title: "SPAG package tutorial (vignette)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{SPAG tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{Cairo}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

# R tool for measuring economic activity

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package provides a method to calculate and visualize the Index of Spatial Agglomeration (SPAG).

## Installation

The package is still in development phase and can be downloaded and installed from GitHub:

```{r eval=FALSE}
devtools::install_github("pbiecek/SPAG")
```

Use the `library()` function to load the package.

```{r, warning=FALSE, message=FALSE}
library("SPAG")
```

## Data sets

In order to calculate the SPAG Index two arguments have to be provided - information about the area and companies for which the index is supposed to be calculated. The area should be provided as a SpatialPolygonsDataFrame. An exemplary map in the format of a SpatialPolygonsDataFrame is provided with the package:

```{r eval=FALSE}
?ShapefilePoland
ShapefilePoland
```

The information about the companies should be provided in the form of a data frame with four columns: the geographical coordinates of the companies (longitude and latitute), the number of people working for the company (numeric) and a categorical value defining the field of work of the company. A data frame with such data is provided in the package:

```{r}
# ?CompaniesPoland
CompaniesPoland
```

## Calculating the index

The package implements function SPAG responsible for calculating the index along with its components. The function takes 4 parameters - the data frame with companies, a spatial data frame with the map and two parameters limiting the number of companies required for calculating the Distance Index.

```{r, message=FALSE, warning=FALSE}
# ?SPAG
SPAGIndex <- SPAG(companiesDF = CompaniesPoland, shp = ShapefilePoland, theoreticalSample=1000, empiricalSample=2000)
print(SPAGIndex)
```

The SPAG function returns an SPAG object: a dataframe containing SPAG index and its components calculated for every category as well as the data required for visualizing the index.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
setwd("C:/Users/Max/Desktop/TestEmpiryczny")
dane<-read.csv("geoloc data.csv", header=TRUE, sep=";", dec=".")
dane$zatr<-ifelse(dane$GR_LPRAC==1, 5, ifelse(dane$GR_LPRAC==2, 30, ifelse(dane$GR_LPRAC==3,150, ifelse(dane$GR_LPRAC==4, 600, 1500))))
CompaniesPoland<-dane[dane$SEK_PKD7 %in% c("B","C","D","E"), c(23,24,25,26) ]
```

The last two parameters - theoreticalSample and empiricalSample limit the number of companies used in the calculation of the Distance Index. This has to be done as the algorithm has time and space complexity of $O(n^2)$. For bigger data sets such algorithm would be too time consuming. Below an example of limiting the calculation for a slightly biger data set is shown. As seen only the distance index changes:

```{r, message=FALSE, warning=FALSE}
startTime <- Sys.time()
SPAGIndexFULL <- SPAG(companiesDF = CompaniesPoland, shp = ShapefilePoland, theoreticalSample=10000, empiricalSample=10000)
endTime <- Sys.time()
print(SPAGIndexFULL)
print(endTime-startTime)

startTime <- Sys.time()
SPAGIndexSAM <- SPAG(companiesDF = CompaniesPoland, shp = ShapefilePoland, theoreticalSample=100, empiricalSample=100)
endTime <- Sys.time()
print(SPAGIndexSAM)
print(endTime-startTime)
```

## Plotting the Index

The plot function implemented for SPAG objects uses `ggplot2` and `ggmap` packages. The function returns an object of class ggplot which can be further modified using functionalities from the aforementioned packages. By default the function returnes the SPAG index calculated for all categories:

```{r}
ggSPAG <- plot(SPAGIndex)
ggSPAG
```

This can be modified by setting the category parameter to match the category for which we want to plot the SPAG Index:

```{r}
plot(SPAGIndex, category="A")
plot(SPAGIndex, category="A", addCompanies = FALSE)
```

By default the index is presented as a map on a white background. This can be changed by adding a theme from ggplot2 package. More information about ggplot2 themes can be found [here](https://www.google.com) :
```{r}
ggSPAG + theme_bw()
```

There are different ways to present a map on a 2 dimensional canvas. By default SPAG Index is plotted using mercator projection but it can be changed by using the `coord_map()` function from package ggplot2:

```{r}
ggSPAG + coord_map("ortho", orientation = c(-10, 5, 0))
ggSPAG + coord_map("conic", lat=20) + coord_flip()
```
